---
title: "Exploratory Data Analysis"
author: "YOUR NAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Reading Notes

2 Key questions: Ask about variation and covariation

Visualization is a good place to start. Categorical or quantitative?

Consider geom_freqpoly for layered graphs

Search for both  typical and unusual values

## EDA Example

The code below is an example of the EDA process using the `starwars` data from the `tidyverse` package. (Make sure you have the `tidyverse` package installed!)

Run the code line-by-line and look at the output. Add a comment to each line of code that explains what it does/what insights it provides.

```{r}
library(tidyverse)
# Adds the tidyverse library to use functions saved there.

dim(starwars) #Gives the dimensions of starwars data
colnames(starwars) #Gives the names of the columns of starwars

head(starwars) #Shows you the first bit of data at the beginning of the tables
glimpse(starwars) #Glimpse is like head but shows the data in a different format

?starwars #Shows the help file about the starwars data

starwars %>% count(species) #Pipes starwars data into the count function which counts the species

mean(starwars$height) #gives the mean of the height column in the starwars data
mean(starwars$height, na.rm = TRUE) #Ignores NA values to do the same as above in spite of NA's included.

summary(starwars$height) #Shows a summary of commonly sought statistics about the starwars height vector

cm_to_ft <- function(cm){
  ft = cm / 30.48
  return(ft)
} #Creates a function that converts cm to ft

starwars_2 <- starwars %>% mutate(height_ft = cm_to_ft(height)) #converts height of starwars characters to ft instead of cm and stors it in a new copy of the dataset. 

dim(starwars_2) #Retrieves dimentions of starwars 2 data
colnames(starwars_2) #Retrieves the column names of starwars 2 data

summary(starwars_2$height_ft) #Shows a summary of the hight data in ft in star wars 2 data

ggplot(starwars_2, aes(height_ft)) + 
  geom_histogram() #creates a histogram of heights in feet from star wars 2 data

starwars_2 %>% filter(height_ft == max(height_ft)) #filters the data for only the very tallest characters, including ties
starwars_2 %>% filter(height_ft == max(height_ft, na.rm = TRUE)) #Does the same but accounts for NA's that would otherwise stop the function from running.

# https://starwars.fandom.com/wiki/Yarael_Poof
```

## EDA Practice

Continue exploring the `starwars` data to gain additional insights, using [R4DS: Chapter 7](https://r4ds.had.co.nz/exploratory-data-analysis.html) as a guide.

It is ok if you don't understand the all code in Chapter 7. (That is what we'll be learning the next two weeks!) If writing your own code is a struggle, try the "copy, paste, and tweak" method.

```{r}
view(starwars)
library(mosaic)
favstats(mass/height ~ sex, data = starwars)
#Here I wanted to compare mass and height for some correlation about density. I divided it by sex because men and women I might expect to have different values. Hermaphroditic characters seem to have very large mass compared to their height.

hist(starwars$mass/starwars$height)
#This histogram helps clarify, there are many characters in the lower range, while only a few at such a high range like the hermaphroditic characters, we can assume that outliers in density match the rare hermaphroditic character type.
```