---
title: "Case Study 4"
author: "Cody Jenson"
date: "2/2/2021"
output: 
  html_document: 
    code_folding: hide
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(nycflights13)
library(tidyverse)
library(mosaic)
library(pander)
```

# Question 1:

### Which origin airport is best to minimize my chances of a late arrival when I am using Delta Airlines?

```{r}
del <- filter(flights, carrier == "DL")
pander(mean(del$arr_delay ~ del$origin, na.rm = TRUE))

ggplot(del) +
  geom_point(aes(x = arr_time, y = sched_arr_time)) +
  facet_grid(cols = vars(origin))
```

This simple table gives the average delay of arrival from each airport for delta planes. Because the delay time is set as a positive number, the negative actually means that delta flights that originate from JFK on average arrive early, so that would be the best airport to choose.

# Question 2:

## Which destination airport is the worst airport for arrival time? (you decide on the metric for "worst")

```{r}
list <- data.frame(mean(flights$arr_delay ~ del$dest, na.rm = TRUE))
list <- list %>%
  rename(average = mean.flights.arr_delay...del.dest..na.rm...TRUE.,)
pander(filter(list, list$average == max((list$average))))
```

STL airport has on average the latest arrival times so that would be the worst airport.

